\LARGE \bf Laser Tracker Placement Optimization for highly flexible manufacturing systems

### 2024-07-29 paper/paper.tex
added marker vis and rephrased experiment section

\author{Jan Baumgärtner$^{1}$,  Alexander Puchta$^{1}$ and Jürgen Fleischer$^{1}$% <-this % stops a space
\thanks{*The authors would like to thank the Ministry of Science, Research and Arts of the Federal State of Baden-Württemberg for the financial support of the project within the InnovationsCampus Future Mobility.
}% <-this % stops a space
\thanks{$^{1}$Jan Baumgärtner, Alexander Puchta, and Jürgen Fleischer are with the wbk Institute of Production Science,
        Karlsruhe Institute of Technology, 76131 Karlsruhe, Germany
        {\tt\small jan.baumgaertner@kit.edu}}%
}

@@

This will naturally force the optimizer to place the marker close to the surface of the manufacturing system.

The pose of the laser tracker is in reality also defined by a transformation matrix $T_{l}$.
However since the laser tracker has a spherical measurment volume we only need to define the position of the laser tracker $p_{p_l}$, which is a three dimensional vector.
The laser tracker is also mounted on a stand, in principle this can vary in height but only in a limited range.
This can be encoded using a simple constraint:
\begin{equation}
    h_{min} \leq p_{p_l,z} \leq h_{max}
\end{equation}
where $h_{min}$ and $h_{max}$ are the minimum and maximum height of the laser tracker stand.
The laser tracker also has a limited range of measurement.
Here we could define a constraint that the marker needs to be within the measurement range of the laser tracker.
But in practice the laser tracker works better the closer the marker is to the laser tracker so long it is in the minimum range.
This can be encoded using a penalty function:
\begin{equation}
    p(p_m,p_l) =  \begin{cases}
        d & \text{if } \text{marker is in measurement range of the laser tracker} \\
        \rho & \text{otherwise}
    \end{cases}
\end{equation}
where $d$ is the distance of the marker to the laser tracker and $\rho$ is a large penalty factor.
The final optimization problem can be formulated as follows:
\begin{equation}
    \begin{split}
    \max_{p_m,p_l} & h(p_m,p_l) \\
    \text{s.t.} & \sqrt{p_{p_m,x}^2 + p_{p_m,y}^2} \leq r \\
    & 0 \leq p_{p_m,z} \leq h \\
    & h_{min} \leq p_{p_l,z} \leq h_{max} \\
    & p(p_m,p_l) \geq 0
    \end{split}
\end{equation}

@@

\begin{table}  %TODO anpassen
        \centering
        \begin{tabular}{c|c}
                Hyperparameter & Value \\
                \hline
                Number of particles & 100 \\
                Number of iterations & 1000 \\
                Inertia weight & 0.9 \\
                Cognitive coefficient & 2.0 \\
                Social coefficient & 2.0 \\
        \end{tabular}
        \caption{Hyperparameters of the PSO algorithm}
        \label{tab:hyperparameters}
\end{table}

@@

\section{Experiment Setup}
To test the performance of our optimiser, we need to define a set of test cases.
These consist of different trajectories of the two robots shown in figure \ref{fig:simulation}.
Both robots are Comau NJ290 industrial robots with a reach of 2.9 metres.
The right robot is used to service the CNC machine and the additive manufacturing system, while the left robot is used to perform a high-precision manufacturing task.
To properly control this task, the left robot needs to be tracked by the laser tracker.
We test whether our algorithm is able to find a placement of the laser tracking system and markers that can track the robot over its entire trajectory.
We compare this to the performance of the original population of marker and laser tracker placements used to start the optimisation.

>>>>

\author{Jan Baumgärtner$^{1}$, Max Goebels,   Alexander Puchta$^{1}$ and Jürgen Fleischer$^{1}$% <-this % stops a space
\thanks{*The authors would like to thank the Ministry of Science, Research and Arts of the Federal State of Baden-Württemberg for the financial support of the project within the InnovationsCampus Future Mobility.
}% <-this % stops a space
\thanks{$^{1}$Jan Baumgärtner, Max Goebels, Alexander Puchta, and Jürgen Fleischer are with the wbk Institute of Production Science,
        Karlsruhe Institute of Technology, 76131 Karlsruhe, Germany
        {\tt\small max.goebels@kit.edu}}%
}

@@

This will naturally force the optimizer to place the marker close to the surface of the manufacturing system.\\
\\
The pose of the laser tracker is in reality also defined by a transformation matrix $T_{l}$.
However since the laser tracker has a spherical measurment volume we only need to define the position of the laser tracker $p_{p_l}$, which is a three dimensional vector.
The laser tracker is also mounted on a stand, in principle this can vary in height but only in a limited range.
This can be encoded using a simple constraint:
    h_{min} \leq p_{p_l,z} \leq h_{max}
where $h_{min}$ and $h_{max}$ are the minimum and maximum height of the laser tracker stand.
The laser tracker also has a limited range of measurement.
Here we could define a constraint that the marker needs to be within the measurement range of the laser tracker.
But in practice the laser tracker works better the closer the marker is to the laser tracker so long it is in the minimum range.
This can be encoded using a penalty function:
    p(p_m,p_l) =  \begin{cases}
        d & \text{if } \text{marker is in measurement range of the laser tracker} \\
        \rho & \text{otherwise}
where $d$ is the distance of the marker to the laser tracker and $\rho$ is a large penalty factor.
The final optimization problem can be formulated as follows:
    \max_{p_m,p_l} & h(p_m,p_l) \\
    \text{s.t.} & \sqrt{p_{p_m,x}^2 + p_{p_m,y}^2} \leq r \\
    & 0 \leq p_{p_m,z} \leq h \\
    & h_{min} \leq p_{p_l,z} \leq h_{max} \\
    & p(p_m,p_l) \geq 0

@@

\begin{table}  %TODO anpassen
        \centering
        \begin{tabular}{c|c}
                Hyperparameter & Value \\
                \hline
                Number of particles & 500 \\
                Number of iterations & 100 \\
                Inertia weight & 0.8 \\
                Cognitive coefficient & 0.1 \\
                Social coefficient & 0.1 \\
        \end{tabular}
        \caption{Hyperparameters of the PSO algorithm}
        \label{tab:hyperparameters}
\end{table}

@@

\section{Experiment Setup}
To test the performance of our optimizer, we need to define a testcase.
Both robots are Comau NJ290 industrial robots with a reach of 2.9 meters.
The right robot services the CNC machine and the additive manufacturing system, while the left robot performs a high-precision manufacturing task.
For this test, the left robot follows a rectangular trajectory with dimensions of 0.4 meters in length and 0.6 meters in width, as depicted in Figure \ref{fig:trajectory}.
The aim is now to find a laser tracker and marker placement that can track the left robot throughout its entire trajectory.
We compare this to the performance of the initial population of marker and laser tracker placements used to start the optimization process.
Figure \ref{fig:iterations} illustrates the performance of the algorithm over multiple iterations.
The results show that the algorithm converges to a solution in roughly 40 iterations, which is relatively fast for an optimization approach.
This speed is likely due to the existence of numerous optimal placements for the laser tracker and markers.
The final solution is also depicted in Figure \ref{fig:trajectory}.

\begin{figure}
\centering
\includegraphics[width=\textwidth]{figures/iterations.pdf}
\caption{Performance of the algorithm over the iterations.}
\label{fig:iterations}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/trajectory.png}
\caption{Trajectory of the robot and final laser tracker and marker placements.}
\label{fig:trajectory}
\end{figure}
